We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-01-31 paper/paper.tex
edit caption

\begin{equation}
p(X; \mu, \Sigma) = \frac{1}{(2\pi)^{3/2}|\Sigma|^{1/2}} \exp\left(-\frac{1}{2}(X-\mu)^{T}\Sigma^{-1}(X-\mu)\right), 
\end{equation}
% 
where $X=[w, c/a, b/a]^{T}$ is a vector variable with the normalized
quantities, $\mu$ is the vector mean and the $\Sigma$ is the
covariance matrix.

@@

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in the Local Group and the
  Isolated Pairs. In all panels the horizontal axis corresponds to the
  M31 or the most massive halo in the pair and the vertical axis to
  the MW or the least massive halo in the pair.
The panel on the left shows the plane width in physical units
comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions while keeping its radial distance (circles). 
The panel in the middle compares the average from the observations
(star) and the average from each one of the Isolated Pairs (circles
with error bars).
The panel on the right has the same information as the middle panel,
only that this time each point has been normalized (median substracted
and normalized by the standar deviation) to the results of its
randomization. 
The main message of this series of plots is that the MW has a
significantly thiner plane both compared to the result of its own
satellite spherical randomization (left panel) and the expectation from
simulations (middle panel). 
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In the M31 their satellites are in agreement with the expectations
both from an spherical distribution and the results form the
simulations. 
A second conclusion is that the spherically averaged plane width
MW (seen in the point cloud in the left panel) is smaller than the
average expectation from simulations, while for M31 the spherical
average is consistent with simulations. 
\label{fig:scatter_width}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ca_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $c/a$ axis ratio. 
The message holds in this case as for the plane width.
The MW is has a significantly low $c/a$ value compared to the
expectation from a spherical distribution and simulations. 
This low value is also $2\sigma$ away from the expactations for an
spherical distribution.
M31 is consistent both with an spherical distribution and the results
from simulations.
However, in this case the axis ratio in the spherically averaged case
is completeley consistent with the expectation from simulations.
\label{fig:scatter_ca_ratio}}
\end{figure*}

@@

The middle panel in \ref{fig:scatter} compares the observational resul
against the measurements of all pairs from the Illustris1-Dark
simulation.
In this case we have a similar result as before. 
The observed MW width is smaller than all the results in the
simulation, there is not a single halo with similar values.
On the other hand, the results for the M31 are entirely consistend
with observations. Most of the halos in the simulation show a width
value similar to M31.

The right panel.

>>>>

\begin{equation}
p(X; \mu, \Sigma) = \frac{1}{(2\pi)^{3/2}|\Sigma|^{1/2}}
\exp\left(-\frac{1}{2}(X-\mu)^{T}\Sigma^{-1}(X-\mu)\right), 
\label{eq:multivariate}
\end{equation}
% 
where $X=[w, c/a, b/a]^{T}$ is a vector variable with the normalized
quantities, $\mu$ is the vector mean and the $\Sigma$ is the
covariance matrix.

@@

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_width.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_width.pdf}
\caption{Plane width characterization in observations and
  simulations (Illustris1-Dark in this case.) In all panels the
  horizontal/vertical axis corresponds to the  M31/MW or the
  most/less massive halo in the pair 
  Left. Plane width in physical units comparing the results from observations
(stars) against the result of spherically randomizing the satellite
positions (circles). 
Middle. Average from the observations (star) and the average from each
pair in the simulation (circles).
Right. Same as the middle panel only that this time each point has
been normalized (median substracted and normalized by the standar
deviation) to the results of its randomization. 
The main message of these panels that the MW has a significantly
thiner plane both compared to the result of its own satellite
spherical randomization (left panel) and the expectation from
simulations (middle and right panel).  
This low value is $2\sigma$ away from what is expected in a spherical
distribution. 
In M31 the satellite distribution are in agreement with the
expectations both from an spherical distribution and the results form the
simulations. 
\label{fig:scatter_width}}
\end{figure*}
\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{scatter_random_ranked_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_ranked_illudm_ca_ratio.pdf}
\includegraphics[width=0.30\textwidth]{scatter_norm_ranked_illudm_ca_ratio.pdf}
\caption{Same layout as in Figure \ref{fig:scatter_width}. 
This time for the $c/a$ axis ratio. 
The message holds in this case as for the plane width.
The MW is has a significantly low $c/a$ value compared to the
expectation from a spherical distribution and simulations. 
This low value is also $2\sigma$ away from the expactations for an
spherical distribution.
M31 is consistent both with an spherical distribution and the results
from simulations.
However, in this case the axis ratio in the spherically averaged case
is completeley consistent with the expectation from simulations.
\label{fig:scatter_ca_ratio}}
\end{figure*}

@@

The middle panel in Figure \ref{fig:scatter_width} compares the
observational result (star) against the measurements of all pairs from the Illustris1-Dark
simulation (circles).
In this case we have a similar result as before. 
The observed MW width is smaller than all the results in the
simulation, there is not a single halo with similar values.
On the other hand, the results for the M31 are entirely consistend
with observations. Most of the halos in the simulation show a width
value similar to M31.

The right panel in Figure \ref{fig:scatter_width} shows the result for
the normalized width.  
This panel tells the same story as the middle panel. The M31 values
are typical while the MW is an outlier. 
The added value of the data in this panel is that it is the normalized
which is consistent with normal distributions.
This is the data used to build the mean values vector and covariance
matrix described in Equation \ref{eq:multivariate}.
